Active Learning from Blackbox to Timed Connectors

### 2016-04-19 shortpaper/paper.tex
.

%\subsection{Observation Table}
%\emph{Observation Tables}, proposed in \cite{DBLP:journals/iandc/Angluin87}, are used to construct
%hypothetical Mealy machines from the results of membership queries.

%Mealy machines consist of \emph{states} and \emph{transitions}.
%As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
%after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
%state of $s$ after executing $a$.
%However, under such notations, there are \emph{infinite} number of states, among which most are
%duplicate.

%States are distinguished, usually, by their \emph{subsequent behavior}. Suppose $s_1\neq s_2\in
%\mathcal{I}^*$, we say the corresponding of states $s_1$ and $s_2$ are equivalent iff $\forall d\in
%\mathcal{I}^+. mq(s_1d)=mq(s_2d)$. Since we can never check it on every $d\in\mathcal{I}^+$, an
%alternative way is to use a set of \emph{suffixes} to distinguish them approximately.
%\begin{definition}[$D$-Equivalence]
  %Provided with a set of suffixes $D\subset\mathcal{I}^+$, the corresponding states of
  %$s_1,s_2\in\mathcal{I}^*$ are $D$-equivalent (denoted as $s_1\sim_D s_2$)
  %iff $mq(s_1d) = mq(s_2d)$ for all $d\in D$.
%\end{definition}

@@

%%\begin{figure}[ht]
  %%\begin{center}
    %%\begin{displaymath}
      %%\begin{array}{l||ccccc||c}
        %%\hline
        %%& (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        %%\hline\hline
        %%\varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        %%\hline
        %%(a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        %%(\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %%\rowcolor[gray]{.9}
        %%(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & unclosed\\
        %%(\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %%T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %%\hline
      %%\end{array}
    %%\end{displaymath}
    %%% \input{./images/hypo.tex}
  %%\end{center}
  %%\caption{Observation Table and Corresponding Hypothesis}
  %%\label{fig:hypo}
%%\end{figure}

@@

\subsection{L* Algorithm}
The whole L* algorithm can be concluded as follows.
\begin{algorithm} 
  \caption{L*} 
  \label{alg:lstar}
  \KwIn{Oracle interfaces $mq,eq$, Input actions $\mathcal{I}$}
  \KwOut{Observation table $obs$} 
  $obs.S$ initialized as empty\;
  $obs.D$ initialized as $\{[i]|i\in\mathcal{I}\}$\;
  \Repeat{$ce=true$}{
    $obs$ = EncloseTable($mq$, $\mathcal{I}$, $obs$)\;
    $ce=eq(obs)$\;
    \If {$ce\neq true$} {
      $obs.D$.append(Suffix$(ce, obs)$)\;
    }
  }
  \Return $obs$\; 
\end{algorithm}

In this algorithm, when we use \emph{observation tables} to represent hypothetical models.

>>>>

%\subsection{Observation Table}
%\emph{Observation Tables}, proposed in \cite{DBLP:journals/iandc/Angluin87}, are used to describe
%hypothesises in form of Mealy machines.
%As mentioned before, Mealy machines are composed of \emph{states} and \emph{transitions}. However, in
%blackboxes, only \emph{input sequences} and their corresponding \emph{outputs} are observable. A
%natural idea is to use input sequences to indicate their corresponding states, and outputs to
%describe the edges.

%\begin{figure}[ht]
  %\begin{center}
    %\footnotesize
    %\begin{displaymath}
      %\begin{array}{l||ccccc}
        %\hline
        %& (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T \\
        %\hline\hline
        %\varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        %\hline
        %(a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty \\
        %(\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty \\
        %\rowcolor[gray]{.9}
        %(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty \\
        %(\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty \\
        %T & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        %\hline
      %\end{array}
    %\end{displaymath}
    %% \input{./images/hypo.tex}
  %\end{center}
  %\caption{Observation Table}
  %\label{fig:hypo}
%\end{figure}

%However, under such notations, there are \emph{infinite} number of states, among which most are
%duplicate.

@@

@@

\subsection{L* Algorithm}
The whole L* algorithm can be briefly concluded as follows. In this algorithm,
\begin{itemize}
  \item \emph{observation table} is a widely-used data structure to describe the hypothesis.
\end{itemize}
Please refer to \cite{DBLP:conf/sfm/SteffenHM11, DBLP:journals/iandc/Angluin87} for further details.
\begin{algorithm} 
  \caption{L*} 
  \label{alg:lstar}
  \KwIn{Oracle interfaces $mq,eq$, Input actions $\mathcal{I}$}
  \KwOut{Observation table $obs$} 
  $obs.S$ initialized as empty\;
  $obs.D$ initialized as $\{[i]|i\in\mathcal{I}\}$\;
  \Repeat{$ce=true$}{
    $obs$ = EncloseHypothesis($mq$, $\mathcal{I}$, $obs$)\;
    $ce=eq(obs)$\;
    \If {$ce\neq true$} {
      $obs.D$.append(Suffix$(ce, obs)$)\;
    }
  }
  \Return $obs$\; 
\end{algorithm}

The L* algorithm takes a set of input actions $\mathcal{I}$ and a oracle interface (which plays the
role of SUL) as its parameters. And it returns a so-called \emph{observation table} as the final
hypothesis. (Observation table is a widely-used data structure, which can represent a Mealy machine
by its external observation.)
In each iteration, the algorithm will:
\begin{enumerate}
  \item \emph{Enclose the hypothesis.} Active learning is very similar to a breadth-first-search
    process on an unknown Mealy machine. At first, only the initial state is visible to us. We then
    explore different input actions and check if they can lead to some state that is already known.
    If every edge targets to an explored state, we will say that the hypothesis is \emph{closed}.
  \item \emph{Equivalence checking.} After the hypothesis is enclosed, we need to check if the
    current hypothesis is equivalent to the real model. Such $eq$ operation returns a
    \emph{true} if they are equivalent, or a counter-example if not.
  \item \emph{Counter-example Analysis.}  
\end{enumerate}
